Curve counting in genus one: elliptic singularities and the splitting formula

### 2019-06-26 Reduced Gromov-Witten Theory of Hypersurfaces in Genus One.tex
removed ipe and included png figures

\subsubsection{Combinatorial types} Recall that a boundary divisor $\Dcal \subseteq \Dcal_{1,\alpha,k}(\PP^N|H,d)$ is said to be of \textbf{type $C^0$} if the core is assigned zero degree by the combinatorial type. In this case, the combinatorial type is not as tightly constrained as previously (it need not be a bipartite graph with a single interior vertex). Nevertheless, we have the following lemmas:
\begin{lemma} 
Let $\Dcal \subseteq \Dcal_{1,\alpha,k}(\PP^N|H,d)$ be a locus of type $C^0$ and let $\Delta$ be the corresponding combinatorial type. Then for every nonzero vertex $\diamond$ of the expanded target, the fibre of the tropical map over $\diamond$ contains exactly one stable vertex (together with a number of trivial bubbles), and at least one vertex which lies on the circle (which may or may not be stable).
\end{lemma}
\begin{proof}
By stability, we always need at least one stable vertex lying over each vertex of the target. If there was more than one, then (since both the insertions and the factorisation condition are pulled back from the moduli space of maps without expansions) it could be seen, as in the proof of Lemma \ref{lem:combs}, that the contribution would vanish.

Furthermore, for each vertex of the target there must always be a vertex in the preimage which lies on the radius, since otherwise there would be at least two independent tropical parameters, and so $\Dcal$ would not be a divisor.\end{proof}

@@

%%%%%%%%
\begin{center}
\begin{figure}
 \begin{tikzpicture}[ipe import]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \draw
    (64, 672)
     -- (128, 688)
     -- (64, 704)
     -- (64, 704);
  \draw
    (128, 688)
     -- (64, 736)
     -- (64, 736);
  \draw[red]
    (128, 688) circle[radius=65.9697];
  \draw
    (128, 688)
     -- (112, 752)
     -- (64, 768)
     -- (64, 768);
  \draw
    (112, 752)
     -- (64, 784)
     -- (64, 784);
  \pic
     at (64, 672) {ipe disk};
  \pic
     at (64, 704) {ipe disk};
  \pic
     at (64, 736) {ipe disk};
  \pic
     at (64, 768) {ipe disk};
  \pic
     at (64, 784) {ipe disk};
  \pic
     at (112, 752) {ipe disk};
  \pic
     at (75.2242, 727.5818) {ipe cross};
  \pic
     at (112, 700) {ipe cross};
  \pic
     at (112, 692) {ipe cross};
  \pic
     at (112, 684) {ipe cross};
  \pic
     at (75.8121, 701.047) {ipe cross};
  \pic
     at (75.0705, 674.4385) {ipe cross};
  \filldraw[fill=white,->]
    (64, 608)
     -- (208, 608);
  \pic
     at (64, 608) {ipe disk};
  \draw
    (128, 688)
     -- (208, 688);
  \pic
     at (73.6978, 608.0203) {ipe cross};
  \pic
     at (111.8758, 608.0203) {ipe cross};
  \pic
     at (127.6736, 608.0203) {ipe cross};
  \pic[fill=white]
     at (128, 688) {ipe fdisk};
  \node[ipe node]
     at (86.863, 690.301) {$m$};
  \node[ipe node]
     at (87.521, 671.212) {$m$};
  \node[ipe node]
     at (48.026, 668.579) {$R_1$};
  \node[ipe node]
     at (48.026, 687.668) {$\ldots$};
  \node[ipe node]
     at (47.368, 700.832) {$R_k$};
  \node[ipe node]
     at (48.026, 725.846) {$R_{k+1}$};
  \node[ipe node]
     at (48.026, 784.429) {$R_s$};
  \node[ipe node]
     at (48.026, 766.657) {$\ldots$};
  \node[ipe node]
     at (108.585, 757.441) {$R_r$};
  \node[ipe node]
     at (80, 744.935) {$\ldots$};
  \node[ipe node]
     at (130.965, 677.794) {$E$};
\end{tikzpicture}
\caption{A ray of type $C^0$.}\label{fig:ray_example}
\end{figure}
\end{center}
%%%%%%%%%%

@@

We write $V_0$ for the collection of stable vertices of the combinatorial type which are mapped to $0$ by the tropical map, and $V_{>0}$ for the stable vertices which are mapped to $\RR_{>0}$. For each vertex $v$ there is associated discrete data $\Gamma_v$, which defines a moduli space of relative stable maps if $v \in V_0$ and a moduli space of rubber maps if $v \in V_{>0}$. Then $\Ecal$ admits a finite splitting morphism onto the fibre product
\begin{equation*} \Ecal \xrightarrow{\rho} \left( \prod_{v \in V_0} \ol\Mcal_{\Gamma_v}(\PP^N|H) \times \prod_{v \in V_{>0}} \ol\Mcal^{\leftrightarrow}_{\Gamma_v}(P|H_0+H_\infty) \right) \times_{H^{2E}} H^E \end{equation*}
where $E$ is the set of edges of the (stabilised) dual graph, and the fibre product is taken over the appropriate evaluation maps. Note that only the factor corresponding to the core vertex has non-zero genus, and that it must have zero degree.

@@

\begin{comment}
In this case, the core has (horizontal) degree $0$ and is mapped into $H$. We start by describing the relevant combinatorial types of rays in $\widetilde{\VZ}_{1,\alpha}(\PP^N|H,d)$.
\begin{lemma} In what follows it is understood that we insert trivial bubbles whenever the curve or target need to be subdivided in order for the image and preimage of every polyhedral subcomplex to be a union of polyhedral subcomplexes.
 \begin{enumerate}
  \item The core is represented by a single vertex of genus one and degree zero, tropically mapped to $\RR_{>0}$.
  \item The vertices adjacent to the core with maximal expansion factor are tropically mapped to $0$.
  \item For every point of the target different from $0$, the fiber of the tropical map contains at most one stable vertex, which either is the core or lies on the circle.
  \item $x_k$ is supported by a vertex tropically mapped to $\RR_{>0}$.
 \end{enumerate}
\end{lemma}
\begin{proof}
 By assumption, the core is mapped to $\RR_>0$, hence the distance of its image from $0$ is one parameter $\ell$; since we are describing a ray, all other parameters have to be related to $\ell$, and the only way it can be so is via the alignment. If the fiber over $c>0$ contains two stable vertices, because both the insertions and the factorisation condition are pulled back from the moduli space without expansions, it can be seen as in Lemma \ref{lem:combs} that the collapsing morphism $\Dcal\to\Fcal$ has positive-dimensional fibers (disconnected rubber), so the corresponding combinatorial type does not contribute to the invariants. We conclude that every stable vertex over $\RR_{>0}$ different from the core has to lie on the circle.
\end{proof}
\begin{center}
\begin{figure}
 \begin{tikzpicture}[ipe import]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \draw
    (64, 672)
     -- (128, 688)
     -- (64, 704)
     -- (64, 704);
  \draw
    (128, 688)
     -- (64, 736)
     -- (64, 736);
  \draw[red]
    (128, 688) circle[radius=65.9697];
  \draw
    (128, 688)
     -- (112, 752)
     -- (64, 768)
     -- (64, 768);
  \draw
    (112, 752)
     -- (64, 784)
     -- (64, 784);
  \pic
     at (64, 672) {ipe disk};
  \pic
     at (64, 704) {ipe disk};
  \pic
     at (64, 736) {ipe disk};
  \pic
     at (64, 768) {ipe disk};
  \pic
     at (64, 784) {ipe disk};
  \pic
     at (112, 752) {ipe disk};
  \pic
     at (75.2242, 727.5818) {ipe cross};
  \pic
     at (112, 700) {ipe cross};
  \pic
     at (112, 692) {ipe cross};
  \pic
     at (112, 684) {ipe cross};
  \pic
     at (75.8121, 701.047) {ipe cross};
  \pic
     at (75.0705, 674.4385) {ipe cross};
  \filldraw[fill=white,->]
    (64, 608)
     -- (208, 608);
  \pic
     at (64, 608) {ipe disk};
  \draw
    (128, 688)
     -- (208, 688);
  \pic
     at (73.6978, 608.0203) {ipe cross};
  \pic
     at (111.8758, 608.0203) {ipe cross};
  \pic
     at (127.6736, 608.0203) {ipe cross};
  \pic[fill=white]
     at (128, 688) {ipe fdisk};
  \node[ipe node]
     at (86.863, 690.301) {$m$};
  \node[ipe node]
     at (87.521, 671.212) {$m$};
  \node[ipe node]
     at (48.026, 668.579) {$R_1$};
  \node[ipe node]
     at (48.026, 687.668) {$\ldots$};
  \node[ipe node]
     at (47.368, 700.832) {$R_k$};
  \node[ipe node]
     at (48.026, 725.846) {$R_{k+1}$};
  \node[ipe node]
     at (48.026, 784.429) {$R_s$};
  \node[ipe node]
     at (48.026, 766.657) {$\ldots$};
  \node[ipe node]
     at (108.585, 757.441) {$R_r$};
  \node[ipe node]
     at (80, 744.935) {$\ldots$};
  \node[ipe node]
     at (130.965, 677.794) {$E$};
\end{tikzpicture}
\caption{A ray of type $C^0$.}\label{fig:ray_example}
\end{figure}
\end{center}

See Figure \ref{fig:ray_example}. We observe that the corresponding stratum $\Ecal$ of $\ol\Mcal_{1,\alpha}(\PP^N|H,d)$ has logarithmic codimension equal to the number of non-core vertices mapping to $\RR_{>0}$: indeed, the combinatorial type of $\Ecal$ can be read off by forgetting the circle, and the logarithmic codimension in $\ol\Mcal_{1,\alpha}(\PP^N|H,d)$ is accounted for by the number of subdivisions of the target. The discrepancy in dimension is to be expected, as the loci of type $C^0$ are contained in the exceptional locus of the blow-up.

The geometry of $\widetilde\Dcal\to\Ecal$ is encoded by the combinatorics of the alignment. The reader should bear the following toric construction in mind: suppose we subdivide a fan, and we want to describe the relation between the varieties corresponding respectively to a cone $\rho$ of the subdivision, and the original cone $\sigma$ containing it. The fan of $X_\tau$ is obtained by projecting the star of $\tau$ to the quotient $N_\tau=N_{\RR}/\langle\tau\rangle_{\RR}$ of the ambient vector space by the linear span of the cone\footnote{We find it useful to encode this within the formalism of extended tropicalisations \cite{?}.}. Notice that the dimension of the vector spaces $N_\rho$ and $N_\sigma$ (or the corresponding toric varieties) is off by the difference $c$ between the dimension of the cones. Furthermore, we may need to subdivide the fan of $X_\sigma$ in order for cones of $\Sigma_\rho$ to be mapped surjectively to cones of $\Sigma_\sigma$. The geometric interpretation of this is that $X_\tau$ is obtained as an appropriate compactification of a toric bundle of dimension $c$ over $X_\sigma$; observe that the fiber dimension may not be constant (but it can be made so by blowing up $X_{\sigma}$, see \cite[Lemma 4.1]{AK}). It is a fundamental insight due to Tevelev \cite{Tevelev} that a good compactification of the toric bundle is suggested by the tropicalisation of the locus we are interested in: here, ``good'' means that the intersection of the interesting locus with the boundary is dimensionally transverse. In the case at hand, the interesting locus is that of maps satisfying the factorisation property (i.e. smoothable maps), and the tropicalisation has been worked out in \cite{RSPW2} to consist of well-spaced maps.

We proceed to describe $\Dcal$ step-by-step, building on $\Ecal$.

\subsubsection{The fiber product} We start from the finite splitting morphism
\begin{multline}\label{eq:product}
 \rho\colon \Ecal \to \left(\prod_{i=1}^k \ol\Mcal_{0,\alpha^{(i)}\cup(m)}(\PP^N|H,d_i) \times \prod_{i=k+1}^s \ol\Mcal_{0,\alpha^{(i)}\cup(m_i)}(\PP^N|H,d_i)\right)\times_{H^s} \\ \prod_{j=k+1}^{r}\ol\Mcal_{0,\alpha^{(j)}\cup(-m_i)_{i \in I_j\subseteq\{k+1,\ldots,s\}}\cup(\mu_j)}^{\leftrightarrow}(P|H_0+H_\infty,d_j)\times_{H^{r-k}} \\
 \ol\Mcal_{1,\alpha^{(0)}\cup(-m,\ldots,-m,-\mu_{k+1},\ldots,-\mu_r)\cup(-m_i)_{i\in\{k+1,\ldots,s\}\setminus(\bigcup I_j)}}^{\leftrightarrow}(P|H_0+H_\infty,0)
\end{multline}
The degree of $\rho$ can be computed as in Lemma \ref{lem:saturation} above. From the logarithmic point of view, $\rho$ is a saturation map; generically it should be thought of as follows. On the fiber product, the tangent vectors of the components on level $0$ (corresponding to the factors in the first line above) are identified with the target tangent line up to finite powers; the saturation is taking roots of these isomorphisms (it is interesting to notice that it fails precisely on Gathmann's boundary comb loci).
\end{comment}

\subsubsection{From $\Ecal$ to $\widetilde{\Dcal}$} The reader should bear the following construction from toric geometry in mind: suppose we subdivide a fan, and we want to describe the relation between the varieties corresponding respectively to a cone $\rho$ of the subdivision, and the original cone $\sigma$ containing it. The fan of $X_\tau$ is obtained by projecting the star of $\tau$ to the quotient $N_\tau=N_{\RR}/\langle\tau\rangle_{\RR}$ of the ambient vector space by the linear span of the cone\footnote{We find it useful to encode this within the formalism of extended tropicalisations \cite{?}.}. Notice that the dimension of the vector spaces $N_\rho$ and $N_\sigma$ (or the corresponding toric varieties) is off by the difference $c$ between the dimension of the cones. The image of $\Sigma_\rho$ provides a subdivision of the fan of $X_\sigma$. The geometric content is that $X_\tau$ is obtained as an appropriate compactification of a toric bundle of dimension $c$ over $X_\sigma$.
%observe that the fiber dimension may not be constant (but it can be made so by blowing up $X_{\sigma}$, see \cite[Lemma 4.1]{AK}).

@@

Let $V_{0,\delta}\subseteq V_0$ be the subset of vertices lying on the circle (besides being sent to $0$ by the tropical map). Notice that they all have the same expansion factor $m$, therefore they are at the same distance from the core, generically on $\Ecal$.\footnote{Indeed, generically, these vertices do not contribute any subdivision of the target, i.e. they do not affect the relative dimension of $\widetilde\Dcal\to\Ecal$.} Another way to see the generic alignment is the following: the normal bundles corresponding to these nodes can be expressed as $\mathbb T_{R_v,q_v}\otimes\mathbb T_{E,q_v}$; for two different vertices, the right-hand factors are canonically identified thanks to the group structure on the elliptic curve, whereas the $m$-th powers of the left-hand factors are canonically identified by comparison with the target tangent line, and $\rho$ can be seen as taking an $m$-th root of this preferred isomophism. On the other hand, the alignment does not persist when the curves degenerate (see Figure \ref{fig:fwise_align}).

\begin{figure}
 \begin{tikzpicture}[ipe stylesheet,scale=.7]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \pic
     at (129.8488, 703.2932) {ipe disk};
  \pic
     at (97.8488, 719.2932) {ipe disk};
  \pic
     at (97.8488, 687.2932) {ipe disk};
  \draw
    (97.8488, 719.2932)
     -- (129.8488, 703.2932)
     -- (97.8488, 687.2932);
  \draw
    (129.8488, 703.2932)
     -- (161.8488, 687.2932);

We modify the boundary of $\Ecal$\footnote{With the log structure obtained by pulling back from Kim's space and modding out by $\mathbb N^{\codim (\sigma)}$, i.e. considering the edges of $E$ to have \emph{infinite length}.} as follows:
\begin{enumerate}
 \item radially aligning the core, and imposing factorisation, thus effectively replacing
 \[\ol\Mcal^{\leftrightarrow}_{1,\mu}(P|H_0+H_\infty)\quad\text{by}\quad\VZ^{\leftrightarrow}_{1,\mu}(P|H_0+H_\infty)\]
 in the product above, with the consequence of introducing a reduced rubber;
 \item requiring that any degeneration of $V_{0,\delta}$ (resp. $\forall v\in V_{>0,\delta}$) remains centrally aligned with respect to the core;\footnote{Notice that this requires us to blow up some diagonal subloci of the product above, see Figure \ref{fig:fwise_align}} this has the effect that on the whole of $\widetilde E$ - including the boundary -, for each $v\in \{V_{0,\delta}\}\cup V_{>0,\delta}$, it is always possible to tell the components of the curve that are at distance $\delta$ from the core along the path denominated $v$, and, if there is more than one for the same $v$, the lengths of these paths are identified, so as to define line bundles $\OO(\delta_v), \forall v\in \{V_{0,\delta}\}\cup V_{>0,\delta}$.
\end{enumerate}
It is clear that $\widetilde \Dcal\to\widetilde\Ecal$ is generically a torus bundle of relative dimension $\codim(\sigma)-1$; the open locus is represented by the moduli space of attachments for the elliptic singularities.
\begin{lemma}
 Generically on $\Ecal$, $\widetilde \Dcal\to\widetilde\Ecal$ is isomorphic to $\mathcal P=\PP_{\widetilde\Ecal}(\bigoplus_{v\in \{V_{0,\delta}\}\cup V_{>0,\delta}}\OO(\delta_v))$.
\end{lemma}
\begin{proof}
 All of the line bundles $\OO(\delta_v)$ share a common factor $\mathbb T_E$, the universal tangent line to the elliptic curve (appropriately twisted by bubbling loci as in \cite[\S 2.3]{VZ}); removing this, when the curve has the prescribed dual graph, $\OO(\delta_v)$ is isomorphic to the tangent line to the rational tail $R_v$ at the gluing node $q_v$ (with the caveat that all those corresponding to $V_{0,\delta}$ are canonically identified). This is the open stratum of the moduli space of attaching data \cite[\SS 2.2-2.3]{SMY2}. On the boundary, the circle of radius $\delta$ will pass through a subset of $\{V_{0,\delta}\}\cup V_{>0,\delta}$, while cutting the complementary subset along the edge between the core and $v$; correspondingly, there is a destabilisation of the curve, and an elliptic singularity containing strictly semistable branches. Tropically, pushing some of the vertices off the circle creates new finite-length edges with coordinates $\ell_v\in\RR_{\geq_0}$ (see Figure \ref{fig:off_we_go}); not all of them can be strictly positive, because the circle needs to pass through at least one vertex (of positive horizontal degree). This is patently the fan of a projective space.
\end{proof}

Notice that $\mathcal P$ is well-defined on the whole of $\widetilde \Ecal$, while $\widetilde \Dcal\to \mathcal P$ is in fact a modification oof the boundary, as dictated by the alignment procedure. In the next section, we extrapolate what we need from this combinatorial setup in order to describe $\Dcal$ explicitly.
\begin{comment}
\subsubsection{The generic alignment} In the first line of the product above, we have grouped the $k$ factors adjacent to the core with maximal expansion factor $m$. It is fundamental to realise that, generically on $\Ecal$, the vertices corresponding to these factors are (equidistant and) already aligned with respect to the core. We can construct a modification $\widetilde\Ecal$ of $\mathcal E$ by:
\begin{enumerate}
 \item radially aligning the core, and imposing the double factorisation condition, thus effectively replacing the factor
 \[\ol\Mcal_{1,\alpha^{(0)}\cup(-m,\ldots,-m,-\mu_{k+1},\ldots,-\mu_r)\cup(-m_i)_{i\in\{k+1,\ldots,s\}\setminus(\bigcup I_j)}}^{\leftrightarrow}(P|H_0+H_\infty,0)\]
 by $\VZ_{1,\alpha^{(0)}\cup(-m,\ldots,-m,-\mu_{k+1},\ldots,-\mu_r)\cup(-m_i)_{i\in\{k+1,\ldots,s\}\setminus(\bigcup I_j)}}^{\leftrightarrow}(P|H_0+H_\infty,0)$ in the product above; in particular, there is a universal cotangent line bundle at the markings $\mathbb L_E$, obtained by suitably twisting any marking cotangent line (they are all isomorphic over a smooth elliptic curve) by the appropriate bubbling loci (see \cite[\S 2.3]{VZ});
 \item aligning the first $k$ factors with respect to the core (this blows up some diagonal subloci of the product, so it is not enough to align the individual factors).
\end{enumerate}
See Figure \ref{fig:fwise_align} below. It is obvious that $\widetilde \Dcal\to \Ecal$ factors through $\widetilde\Ecal$. On $\widetilde\Ecal$, we have a tropical line bundle $\OO(\delta)$ representing the minimum distance from the core to the tropical preimage of $0$; generically, this line bundle coincides with the normal bundle of the locus where the node between the core and one (any) of the first $k$ vertices persists, but on the boundary this needs to be twisted by further smoothing parameters according to the (partial) alignment on $\widetilde\Ecal$. We can obtain a line bundle $\mathbb T_R$ accounting only for the behaviour of the rational side of the fiber product by twisting $\OO(\delta)$ with $\mathbb L_E$.
\begin{figure}
 \begin{tikzpicture}[ipe stylesheet,scale=.7]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \pic
     at (129.8488, 703.2932) {ipe disk};
  \pic
     at (97.8488, 719.2932) {ipe disk};
  \pic
     at (97.8488, 687.2932) {ipe disk};
  \draw
    (97.8488, 719.2932)
     -- (129.8488, 703.2932)
     -- (97.8488, 687.2932);
  \draw
    (129.8488, 703.2932)
     -- (161.8488, 687.2932);

@@

\begin{figure}
\begin{tikzpicture}[ipe import]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \draw
    (64, 720)
     -- (112, 736)
     -- (96, 752);
  \draw[->]
    (64, 704)
     -- (128, 704);
  \pic
     at (64, 704) {ipe disk};
  \pic
     at (64, 720) {ipe disk};
  \pic
     at (96, 752) {ipe disk};
  \draw[red, dotted]
    (60.6969, 708.808)
     arc[start angle=-161.5651, end angle=-71.5651, x radius=53.2078, y radius=-65.2105];
  \pic[fill=white]
     at (112, 736) {ipe fdisk};
  \draw
    (192, 720)
     -- (240, 736)
     -- (224, 752);
  \draw[->]
    (192, 704)
     -- (256, 704);
  \pic
     at (192, 704) {ipe disk};
  \pic
     at (192, 720) {ipe disk};
  \pic
     at (224, 752) {ipe disk};
  \pic[fill=white]
     at (240, 736) {ipe fdisk};
  \draw
    (112, 736)
     -- (128, 736);
  \draw
    (240, 736)
     -- (256, 736);
  \draw
    (208.6374, 743.6181)
     -- (224.4352, 750.8588);
  \draw
    (208.6374, 758.7577)
     -- (225.0934, 751.517);
  \pic
     at (207.9791, 742.9599) {ipe disk};
  \pic
     at (207.9791, 758.7577) {ipe disk};
  \pic
     at (208, 704) {ipe cross};
  \pic
     at (224, 704) {ipe cross};
  \pic
     at (240, 704) {ipe cross};
  \pic
     at (96, 704) {ipe cross};
  \pic
     at (112, 704) {ipe cross};
  \draw[red, dotted]
    (231.0126, 755.9018)
     { [rotate=-17.595] arc[start angle=135, end angle=225, x radius=21.7093, y radius=22.6274] };
  \draw[red]
    (188.9249, 712.0225)
     arc[start angle=-156.5092, end angle=-115.3692, x radius=85.088, y radius=-96.462];
\end{tikzpicture}
\caption{On the left, a generic point of $\mathcal P$. On the right, a point of $\Xi$. The dotted circle is the one we draw on $\mathcal P$. The continuous circle represents one of the possible choices of $\delta$ enlarging the dotted one (a stratum of $\widetilde\Dcal$).}\label{fig:principalisation}
\end{figure}

>>>>

\subsubsection{Combinatorial types} Recall that a boundary divisor $\Dcal \subseteq \Dcal_{1,\alpha,k}(\PP^N|H,d)$ is said to be of \textbf{type $C^0$} if the core is assigned zero degree by the combinatorial type. In this case, the combinatorial type is not as tightly constrained as previously (it needs not be a bipartite graph with a single interior vertex). Nevertheless, we have the following lemmas:
\begin{lemma} 
Let $\Dcal \subseteq \Dcal_{1,\alpha,k}(\PP^N|H,d)$ be of type $C^0$ and let $\Delta$ be the corresponding combinatorial type. Then, for every vertex $\diamond\neq0$ of the expanded target, the fibre of the tropical map over $\diamond$ contains exactly one stable vertex (together with a number of trivial bubbles), and at least one vertex which lies on the circle (which may or may not be stable).
\end{lemma}
\begin{proof}
By stability, we always need at least one stable vertex lying over each vertex of the target. If there was more than one, then (since both the insertions and the factorisation condition are pulled back from the moduli space of maps without expansions) it could be seen, as in the proof of Lemma \ref{lem:combs}, that the contribution would vanish.

Furthermore, for each vertex of the target there must always be a vertex in the preimage which lies on the circle, since otherwise there would be at least two independent tropical parameters, and so $\Dcal$ would not be a divisor.\end{proof}

@@

%%%%%%%%
\begin{center}
\begin{figure}
\includegraphics[scale=.6]{ray_example}
\caption{A ray of type $C^0$.}\label{fig:ray_example}
\end{figure}
\end{center}
%%%%%%%%%%

@@

We write $V_0$ for the collection of stable vertices of the combinatorial type which are mapped to $0$ by the tropical map, and $V_{>0}$ for the stable vertices which are mapped to $\RR_{>0}$. For each vertex $v$ there is associated discrete data $\Gamma_v$, which defines a moduli space of relative stable maps if $v \in V_0$ and a moduli space of rubber maps if $v \in V_{>0}$. Then $\Ecal$ admits a finite splitting morphism onto the fibre product
\begin{equation*} \Ecal \xrightarrow{\rho} \left( \prod_{v \in V_0} \ol\Mcal_{\Gamma_v}(\PP^N|H) \times \prod_{v \in V_{>0}} \ol\Mcal^{\leftrightarrow}_{\Gamma_v}(P|H_0+H_\infty) \right) \times_{H^{2E}} H^E \end{equation*}
where $E$ is the set of edges of the (stabilised) dual graph, and the fibre product is taken over the appropriate evaluation maps. Note that only the factor corresponding to the core vertex has non-zero genus, and that it has zero degree by assumption.

@@

\subsubsection{From $\Ecal$ to $\widetilde{\Dcal}$} The reader should bear the following construction from toric geometry in mind: suppose we subdivide a fan, and we want to describe the relation between the varieties corresponding to a cone $\rho$ of the subdivision, and to the original cone $\sigma$ containing it, respectively. For any cone $\tau$, the fan of $X_\tau$ is obtained by projecting the star of $\tau$ to the quotient $N_\tau=N_{\RR}/\langle\tau\rangle_{\RR}$ of the ambient vector space by the linear span of $\tau$.\footnote{This can be encoded in the formalism of extended tropicalisations, see \cite{ACP} and references therein.} Notice that the dimension of the vector spaces $N_\rho$ and $N_\sigma$ (or the corresponding toric varieties) is off by the difference $c$ between the dimension of the cones. The image of $\Sigma_\rho$ provides a subdivision of $\Sigma_\sigma$. The geometric content is that $X_\rho$ is obtained as an appropriate compactification of a toric bundle of dimension $c$ over $X_\sigma$.
%observe that the fiber dimension may not be constant (but it can be made so by blowing up $X_{\sigma}$, see \cite[Lemma 4.1]{AK}).

@@

Let $V_{0,\delta}\subseteq V_0$ be the subset of vertices lying on the circle (besides being sent to $0$ by the tropical map). Notice that they all have the same expansion factor $m$, therefore they are at the same distance from the core, generically on $\Ecal$.\footnote{Indeed, generically, these vertices do not produce any subdivision of the target, so they contribute to the relative dimension of $\widetilde\Dcal\to\Ecal$ as one.} Another way to see the generic alignment is the following: the normal bundles corresponding to these nodes can be expressed as $\mathbb T_{R_v,q_v}\otimes\mathbb T_{E,q_v}$; for two different vertices, the right-hand factors are canonically identified thanks to the group structure on the elliptic curve, whereas the $m$-th powers of the left-hand factors are canonically identified by comparison with the target tangent line, and $\rho$ can be seen as taking an $m$-th root of this preferred isomophism. On the other hand, the alignment does not persist when the curves degenerate (see Figure \ref{fig:fwise_align}).

\begin{figure}
\includegraphics[scale=.5]{factorwise_alignment}
\caption{Two degenerations of the combinatorial type above, which represents a ray. On the left-hand side, the alignment holds only generically. On the right-hand side, it is not enough to align each individual factor.}\label{fig:fwise_align}
\end{figure}

We modify $\Ecal$\footnote{With the log structure obtained by pulling back from Kim's space and modding out by $\mathbb N^{\codim (\sigma)}$, i.e. considering the edges of $E$ to have \emph{infinite length}.} along the boundary as follows:
\begin{enumerate}[leftmargin=.5 cm]
 \item radially aligning the core, and imposing fiberwise factorisation, thus replacing
 \[\ol\Mcal^{\leftrightarrow}_{1,\mu}(P|H_0+H_\infty,d=0)\quad\text{by}\quad\VZ^{\leftrightarrow}_{1,\mu}(P|H_0+H_\infty,d=0)\]
 in the product above, thus introducing a reduced rubber with point target;
 \item requiring that any degeneration of $V_{0,\delta}$ (resp. $\forall v\in V_{>0,\delta}$) remains centrally aligned with respect to the core;\footnote{Notice that this requires us to blow up some diagonal subloci of the product above, see Figure \ref{fig:fwise_align}} as a consequence, on the whole of $\widetilde \Ecal$ - including the boundary -, for each $v\in \{V_{0,\delta}\}\cup V_{>0,\delta}$, it is always possible to tell the components of the curve that are at distance $\delta$ from the core along the path denominated $v$, and, if there is more than one for the same $v$, the lengths of these paths are identified, so as to define line bundles $\OO(\delta_v), \forall v\in \{V_{0,\delta}\}\cup V_{>0,\delta}$.
\end{enumerate}
It is clear that $\widetilde \Dcal\to\widetilde\Ecal$ is generically a torus bundle of relative dimension $\codim(\sigma)-1$; the open locus is represented by the moduli space of attachments for the elliptic singularities.
\begin{lemma}
 Generically on $\Ecal$, $\widetilde \Dcal\to\widetilde\Ecal$ is isomorphic to $\mathcal P=\PP_{\widetilde\Ecal}(\bigoplus_{v\in \{V_{0,\delta}\}\cup V_{>0,\delta}}\OO(\delta_v))$.
\end{lemma}
\begin{proof}
 All of the line bundles $\OO(\delta_v)$ share a common factor $\mathbb T_E$, the universal tangent line to the elliptic curve (appropriately twisted by bubbling loci as in \cite[\S 2.3]{VZ}); removing this, when the curve has the prescribed dual graph, $\OO(\delta_v)$ is isomorphic to the tangent line to the rational tail $R_v$ at the gluing node $q_v$ (with the caveat that all those corresponding to $V_{0,\delta}$ are canonically identified). This is the open stratum of the moduli space of attaching data \cite[\S\S 2.2-2.3]{SMY2}. On the fiber boundary, the circle of radius $\delta$ will pass through a subset of $\{V_{0,\delta}\}\cup V_{>0,\delta}$, while cutting the complementary subset along the edge between the core and $v$; correspondingly, there is a destabilisation of the curve, and an elliptic singularity containing strictly semistable branches. Tropically, pushing some of the vertices off the circle creates new finite-length edges with coordinates $\ell_v\in\RR_{\geq_0}$ (see Figure \ref{fig:off_we_go}); not all of them can be strictly positive, because the circle needs to pass through at least one vertex (of positive horizontal degree). This is patently the fan of a projective space.
\end{proof}
\begin{figure}
\includegraphics[scale=.3]{off_we_go}
\caption{Moving vertices off the circle creates new tropical parameters.}\label{fig:off_we_go}
\end{figure}

Notice that $\mathcal P$ is well-defined on the whole of $\widetilde \Ecal$, while $\widetilde \Dcal\to \mathcal P$ is in fact a modification along the boundary, as dictated by the alignment procedure. In the next section, we extrapolate what we need from this combinatorial setup in order to describe $\Dcal$ explicitly.
\begin{comment}
\subsubsection{The generic alignment} In the first line of the product above, we have grouped the $k$ factors adjacent to the core with maximal expansion factor $m$. It is fundamental to realise that, generically on $\Ecal$, the vertices corresponding to these factors are (equidistant and) already aligned with respect to the core. We can construct a modification $\widetilde\Ecal$ of $\mathcal E$ by:
\begin{enumerate}
 \item radially aligning the core, and imposing the double factorisation condition, thus effectively replacing the factor
 \[\ol\Mcal_{1,\alpha^{(0)}\cup(-m,\ldots,-m,-\mu_{k+1},\ldots,-\mu_r)\cup(-m_i)_{i\in\{k+1,\ldots,s\}\setminus(\bigcup I_j)}}^{\leftrightarrow}(P|H_0+H_\infty,0)\]
 by $\VZ_{1,\alpha^{(0)}\cup(-m,\ldots,-m,-\mu_{k+1},\ldots,-\mu_r)\cup(-m_i)_{i\in\{k+1,\ldots,s\}\setminus(\bigcup I_j)}}^{\leftrightarrow}(P|H_0+H_\infty,0)$ in the product above; in particular, there is a universal cotangent line bundle at the markings $\mathbb L_E$, obtained by suitably twisting any marking cotangent line (they are all isomorphic over a smooth elliptic curve) by the appropriate bubbling loci (see \cite[\S 2.3]{VZ});
 \item aligning the first $k$ factors with respect to the core (this blows up some diagonal subloci of the product, so it is not enough to align the individual factors).
\end{enumerate}
See Figure \ref{fig:fwise_align} below. It is obvious that $\widetilde \Dcal\to \Ecal$ factors through $\widetilde\Ecal$. On $\widetilde\Ecal$, we have a tropical line bundle $\OO(\delta)$ representing the minimum distance from the core to the tropical preimage of $0$; generically, this line bundle coincides with the normal bundle of the locus where the node between the core and one (any) of the first $k$ vertices persists, but on the boundary this needs to be twisted by further smoothing parameters according to the (partial) alignment on $\widetilde\Ecal$. We can obtain a line bundle $\mathbb T_R$ accounting only for the behaviour of the rational side of the fiber product by twisting $\OO(\delta)$ with $\mathbb L_E$.
\begin{figure}
 \begin{tikzpicture}[ipe stylesheet,scale=.7]
  %\useasboundingbox (0, 0) rectangle (595, 842);
  \pic
     at (129.8488, 703.2932) {ipe disk};
  \pic
     at (97.8488, 719.2932) {ipe disk};
  \pic
     at (97.8488, 687.2932) {ipe disk};
  \draw
    (97.8488, 719.2932)
     -- (129.8488, 703.2932)
     -- (97.8488, 687.2932);
  \draw
    (129.8488, 703.2932)
     -- (161.8488, 687.2932);

@@

\begin{figure}
\includegraphics[scale=.5]{principalisation}
\caption{On the left, a generic point of $\mathcal P$. On the right, a point of $\Xi$. The dotted circle is the one we draw on $\mathcal P$. The continuous circle represents one of the possible choices of $\delta$ enlarging the dotted one (a stratum of $\widetilde\Dcal$).}\label{fig:principalisation}
\end{figure}
